### 2023-03-05 main/chapter_theory.tex
finish particle properties table

\section{The Standard Model}
\label{sec:theory_sm}
The current Standard Model features fermions
(of four types in three generations),
which interact through forces mediated by vector bosons
and the scalar Higgs boson.
Massless gluons $g$ mediate the strong force,
the massless photon $\gamma$ mediates the electromagnetic force,
and very massive $W^+$, $W^-$, and $Z$ bosons mediate the weak force.
These fundamental particles of the Standard Model are summarized in
Table~\ref{tab:theory_particles_sm}.

@@

                            & Mass$~\eV[G]$                       & $Q_{EM}$             & $Y_L$                & $T^3_L$              & $Y_R$                & $T^3_R$              \\
\multicolumn{1}{l}{Leptons} & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\

@@

\multicolumn{1}{l}{Quarks}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\

@@

$t$                         & $172.7\pm0.3$                       & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
$b$                         & $4.18^{+0.03}_{-0.02}$              & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\
\multicolumn{1}{l}{Bosons}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\
$g$                         & $0$                                 & $0$                  &                      &                      &                      &                      \\
$W^+$                       & $80.4$                              & $+1$                 &                      &                      &                      &                      \\
$W^-$                       & $80.4$                              & $-1$                 &                      &                      &                      &                      \\
$Z$                         & $91.2$                              & $0$                  &                      &                      &                      &                      \\
$\gamma$                    & $0$                                 & $0$                  &                      &                      &                      &                      \\
$h$                         & $125.3 \pm 0.2$                     & $0$                  &                      &                      &                      &

@@

Properties of Standard Model particles.

@@

% TODO reference for properties: textbooks?

>>>>

\section{The Standard Model}
\label{sec:theory_sm}
The current Standard Model features fermions
(of four types in three generations),
which interact through forces mediated by vector bosons
and the scalar Higgs boson.
Massless gluons $g$ mediate the strong force,
the massless photon $\gamma$ mediates the electromagnetic force,
and very massive $W^+$, $W^-$, and $Z$ bosons mediate the weak force.
These fundamental particles of the Standard Model are summarized in
Table~\ref{tab:theory_particles_sm}.
The masses and charges of these particles are further detailed in
Table~\ref{tab:theory_particles_sm_properties}.

@@

Fermions                    & Mass$\,\eV[G]$                       & $Q_{\mathrm{EM}}$             & $Y_L$                & $T^3_L$              & $Y_R$                & $T^3_R$              \\[0.2em]
\multicolumn{1}{l}{Leptons} & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\[0.2em]

@@

$\nu_\mu$                   & $\lesssim 1.1\times10^{-9}$         & $0$                  & $-\frac{1}{2}$       & $+\frac{1}{2}$       & $0$                  & $0$                  \\[1em]
\multicolumn{1}{l}{Quarks}  & \multicolumn{1}{l}{}                & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} & \multicolumn{1}{l}{} \\[0.2em]

@@

$t$                         & $172.69\pm0.30$                     & $+\frac{2}{3}$       & $+\frac{1}{6}$       & $+\frac{1}{2}$       & $+\frac{2}{3}$       & $0$                  \\
$b$                         & $4.18^{+0.03}_{-0.02}$              & $-\frac{1}{3}$       & $+\frac{1}{6}$       & $-\frac{1}{2}$       & $-\frac{1}{3}$       & $0$                  \\[1em]
\multicolumn{1}{l}{Bosons}  & Mass$\,\eV[G]$                       & $Q_{\mathrm{EM}}$             & $Y$                  & $T^3$                &                      &                      \\[0.2em]
$g$                         & $0$                                 & $0$                  & $0$                  & $0$                  &                      &                      \\
$W^+$                       & $80.4$                              & $+1$                 & $0$                  & $+1$                 &                      &                      \\
$W^-$                       & $80.4$                              & $-1$                 & $0$                  & $-1$                 &                      &                      \\
$Z$                         & $91.2$                              & $0$                  & $0$                  & $0$                  &                      &                      \\
$\gamma$                    & $0$                                 & $0$                  & $0$                  & $0$                  &                      &                      \\
$h$                         & $125.3 \pm 0.2$                     & $0$                  & $+\frac{1}{2}$       & $-\frac{1}{2}$       &                      &

@@

Properties of Standard Model particles~\cite{thomson2013modern,pdg2022ynf}.

@@

% electric charge
The electric charge is given by $Q_{\mathrm{EM}} = Y + T^3$; one can check
that the left (L) and right (R) -handed fermions have equal electric charges
corresponding to their left and right weak hypercharges ($Y_L$ and $Y_R$)
and third weak isospin components ($T^3_L$ and $T^3_R$).
